---
title: "Import Excel"
author: "Yiðit Çetinel"
date: "12 Nisan 2017"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
europe <- read_excel("europe.xlsx")
```

## How to import Excel file into R

1. Use "Import Dataset" under Environment tab
2. Click "From Excel"
3. Click "Browse" and select "europe.xlsx" from where you locate it
4. Change name as "europe" and click "Import"

```{r}
View(europe)
```

Check the column names:

```{r}
names(europe)
```

Second one and fourth one look complicated, so we can change the column names:

```{r}
names(europe)[2] <- "Area"
names(europe)[4] <- "Density"
```

Check the head of the data now:

```{r}
head(europe)
```

Let us plot Area vs Population of the europe data:

```{r}
plot(Area ~ Population, europe, col="red", pch=16, cex=.6)
text(Area ~ Population, europe, labels=Name, cex=.6)
```

Linear regression using Area and Population:

```{r}
model <- lm(Area ~ Population, europe)
plot(Area ~ Population, europe, col="red")
abline(model, col="blue")
```

The model is not good because of the last data (Russia). Let us try again without it.

```{r}
europeWithoutRussia <- subset(europe, Name!="Russia")
model <- lm(Area ~ Population, europeWithoutRussia)
plot(Area ~ Population, europeWithoutRussia, col="red",pch=16,cex=.6)
text(Area ~ Population, europeWithoutRussia, labels=Name,cex=.6)
abline(model, col="blue")
```